Light inextensible strings under tension in the Schwarzschild geometry

### 2019-08-27 string_physics_teacher.tex
more work

General relativity is a topic of long-standing fascination for
undergraduate physics students [CITATION NEEDED].  Einstein's concept
of ``curved'' spacetime is an enduring metaphor that attracts large
numbers of undergraduates into introductory general relativity.

Here I show how numerical methods, including computer algebra systems,
can be used to interpret the concept of ``straight line'' in the
vicinity of a black hole.  The problem is would be suitable for an
undergraduate homework problem~\cite{ajpeditor}.

A light string's being under tension means that its path will be an
extremum of proper length: the calculus of variations proves that the
path is straight in the sense of zero curvature.  In terrestrial
gravity, standard $25\times 10^{-6}\mbox{kg}\cdot\mbox{m}^{-1}$ cotton
sewing thread at its breaking strain of about $40\,\mbox{N}$ would
have a sag of~$\sim 6\times 10^{-6}\mbox{m}$ over a $1\,\mbox{m}$
horizontal scale: barely perceptible.  How does such a quotidian
observation translate into the vicinity of a black hole?  Would the
``warping of space'' result in {\em visibly} curved thread?

@@

If, conversely, $r'(\phi)\neq 0$ whenever~$r>2M$, then the string has
only one free end; the other is tangential to the event horizon (in
the sense that ${\displaystyle \lim_{r\longrightarrow 1^+}r'=0}$),
except in the degenerate case of constant~$\phi$ when the string is
radial for~$r>2M$.

>>>>

General relativity is a topic of long-standing fascination for
undergraduate physics students~\cite{christensen2012}.  Here I show
how numerical methods, including computer algebra systems, can be used
to interpret the concept of ``straight line'' in the vicinity of a
black hole.  The analysis would be suitable for an undergraduate
homework problem~\cite{romano2019}.

A light string's being under tension means that its path will be an
extremum of proper length: the calculus of variations proves that the
path is straight in the sense of zero curvature.  In terrestrial
gravity, standard $25\times 10^{-6}\unit{kg}\cdot\meter^{-1}$ cotton
sewing thread at its breaking strain of about $40\unit{N}$ would have
a sag of~$\sim 6\times 10^{-6}\meter$ over a $1\meter$ horizontal
scale: barely perceptible.  How does such a quotidian observation
translate into the vicinity of a black hole?  Would the ``warping of
space'' result in {\em visibly} curved thread?

@@

If, conversely, $r'(\phi)\neq 0$ whenever~$r>2M$, then the string has
only one free end; the other is tangential to the event horizon (in
the sense that ${\displaystyle \lim_{r\longrightarrow 2M^+}r'=0}$),
except in the degenerate case of constant~$\phi$ when the string is
radial for~$r>2M$.
